### 2024-02-07 si/confidenceIntervals.Rnw
minor edits to response, added H2 profiles to si

\documentclass[11pt]{article}

\input{inputs/header}

\setcounter{tocdepth}{1}

<<packages, include=FALSE>>=
library(tidyverse)
library(pomp)
library(haitipkg)
library(spatPomp)
library(doParallel)
library(doRNG)
library(knitr)
library(latex2exp)

options(
  scipen = 2,
  help_type = "html",
  stringsAsFactors = FALSE,
  continue = "+  ",
  width = 70,
  useFancyQuotes = FALSE,
  reindent.spaces = 2,
  xtable.comment = FALSE
)

opts_chunk$set(
  tidy.opts=list(
    keep.blank.line=FALSE
  ),
  comment="",
  fig.path = 'figure/',
  echo=FALSE,
  fig.align="center",
  dev="pdf",
  dev.args=list(
    bg="transparent",
    pointsize=9
  )
)

if (!dir.exists("../model3/si")) {
  dir.create("../model3/si", recursive = TRUE)
}
@

\begin{document}

%%%%%%%%%% START

\section{Confidence Intervals for Model Parameters}\label{sec:ci}

In this section we provide confidence intervals for all model parameters, excluding those that take unique values for each spatial unit.
For each model and parameter, we use profile log-likelihoods to obtain confidence intervals. TODO: Cite.
Due to the non-linear and stochastic nature of Models~1 and 3, exact evaluations of the profile log-likelihood are difficult to obtain.
Instead, the log-likelihood at each point of the profile is estimated using via Monte-Carlo based particle filter methods.
We therefore obtain confidence intervals for the parameters of Model~1 and Model~3 using the Monte Carlo adjust profile (MCAP) algorithm. TODO: Cite.

\subsection{Model~1 parameters}

<<Load Model 1 Confidence Intervals, include=FALSE, message=FALSE, echo=FALSE>>=
H1_profiles <- readRDS("../model1/run_level_3/h1_profiles.rds")

# TAKEN FROM STACK OVERFLOW:
# https://stackoverflow.com/a/44958156
midcut<-function(x,from,to,by) {
  ## cut the data into bins...
  x=cut(x,seq(from,to,by),include.lowest=T)
  ## make a named vector of the midpoints, names=binnames
  vec=seq(from+by/2,to-by/2,by)
  names(vec)=levels(x)
  ## use the vector to map the names of the bins to the midpoint values
  unname(vec[x])
}

prof_params <- c(
  'betat', 'tau_epi', 'tau_end', 'rho', 'nu', 'sig_sq_epi',
  'sig_sq_end', 'E_0', 'I_0', 'beta1', 'beta2', 'beta3',
  'beta4', 'beta5', 'beta6'
)

H1_profiles_long <- H1_profiles %>%
  select(prof_var, logLik, logLik_se, all_of(prof_params), -E_0, -I_0, -tau_end, -tau_epi, -rho) %>%
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(variable == prof_var) %>%
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 1)

# Tau done on log-scale
H1_long_tau <- H1_profiles %>%
  select(prof_var, logLik, logLik_se, tau_end, tau_epi) %>%
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(value <= 1e+12) %>%
  mutate(value = log(value)) %>%
  filter(variable == prof_var) %>%
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 1)

H1_long_rho <- H1_profiles %>%
  select(prof_var, logLik, logLik_se, rho) %>%
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(value > 0.15) %>%
  filter(variable == prof_var) %>%
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 1)

# E_0, I_0 need to be treated slightly differently
H1_E0_I0_long <- H1_profiles %>%
  select(prof_var, logLik, logLik_se, E_0, I_0) %>%
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(variable == prof_var) %>%
  mutate(value = value * 10911819) %>%
  mutate(value = midcut(value, from = -499, to = 24501, by = 1000)) %>%
  # mutate(value = value / 10911819) %>% 
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 1)

H1_profiles_long <- bind_rows(H1_profiles_long, H1_E0_I0_long, H1_long_tau, H1_long_rho)
@

<<Create H1 Profile CIs>>=
all_mcaps <- list()
all_ci <- list()
for (p in prof_params) {

  mcap_tmp <- mcap(
    logLik = H1_profiles_long %>% filter(prof_var == p) %>% pull(logLik),
    parameter = H1_profiles_long %>% filter(prof_var == p) %>% pull(value)
  )

  all_mcaps[[p]] <- mcap_tmp
  tmp_ci <- mcap_tmp$ci
  names(tmp_ci) <- c("lower", "upper")

  all_ci[[p]] <- c(tmp_ci, 'mle' = mcap_tmp$mle)
}

all_mcaps <- purrr::map_df(
  prof_params,
  function(x) all_mcaps[[x]]$fit %>% mutate(prof_var = x)
)

all_ci <- all_ci |> as.data.frame() |> t() |> as.data.frame()
all_ci$prof_var <- rownames(all_ci)

H1_profiles %>% slice_max(order_by = logLik) %>% 
  select(logLik, all_of(prof_params)) %>% 
  mutate(E_0 = E_0 * 10911819, I_0 = I_0 * 10911819,
         tau_epi = log(tau_epi), tau_end = log(tau_end)) %>%
  pivot_longer(
    cols = -logLik,
    names_to = 'prof_var',
    values_to = 'value'
  ) -> mles
@

\begin{figure}[ht]
<<m1Profs, fig.height=5, fig.width=6, echo=FALSE, warning=FALSE, message=FALSE>>=
# Create a mapping between department names and what should be shown in figures
h1_plots <- list()
for (i in 1:length(prof_params)) {
  p = prof_params[i]
  if (p == 'betat') {
    my_lab <- TeX('$\\zeta')
  } else if (grepl('^beta[[:digit:]]$', p)) {
    num <- gsub('^beta', '', p)
    my_lab <- TeX(paste0('$\\beta_', num, '$'))
  } else if (p == 'E_0') {
    my_lab <- TeX('$E_0(0)$')
  } else if (p == 'I_0') {
    my_lab <- TeX('$I_0(0)$')
  } else if (p == 'nu' || p == 'rho') {
    my_lab <- TeX(paste0('$\\', p, '$'))
  } else if (p == 'sig_sq_epi') {
    my_lab <- TeX('$\\sigma$: Epidemic')
  } else if (p == 'sig_sq_end') {
    my_lab <- TeX('$\\sigma$: Endemic')
  } else if (p == 'tau_end') {
    my_lab <- TeX('$\\log(\\psi)$: Endemic')
  } else if (p == 'tau_epi') {
    my_lab <- TeX('$\\log(\\psi)$: Epidemic')
  }

  gg_tmp <- ggplot() +
    geom_point(
      data = H1_profiles_long %>% filter(prof_var == p), 
      aes(x = value, y = logLik)
    ) +
    geom_line(
      data = all_mcaps %>% filter(prof_var == p), 
      aes(x = parameter, y = smoothed), 
      col = 'blue'
    ) +
    # geom_line(data = all_mcaps, aes(x = parameter, y = quadratic), col = 'red') +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = lower), linetype = 'dashed'
    ) +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = upper), 
      linetype = 'dashed'
    ) +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = mle), col = 'blue'
    ) +
    # This code chunk below will plot the MLE along with the smoothed marginal 
    # MLEs and the confidence interval. 
    # geom_point(
    #   data = mles %>% filter(prof_var == p), 
    #   aes(x = value, y = logLik), col = 'red'
    # ) +
    labs(y = 'Log-Likelihood', title = my_lab) +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
    ) +
    theme_bw() +
    scale_y_continuous(limits = c(-2770, max(H1_profiles$logLik))) + 
    theme(
      axis.title.x = element_blank(), 
      plot.title = element_text(
        hjust = 0.5, size = 8, margin = margin(t = 0, unit = "pt")
        ),
      axis.title.y = element_text(size = 8),
      axis.text.y = element_text(size = 7),
      axis.text.x = element_text(size = 6.5),
      plot.margin = margin(5.5, 3, 0, -1, unit = 'pt')
    )

  if (p == 'E_0') {
    gg_tmp <- gg_tmp + 
      scale_x_continuous(breaks = c(0, round(all_ci[p, 'upper'])))
  } else if (p == 'I_0') {
    gg_tmp <- gg_tmp + 
      scale_x_continuous(
        breaks = c(round(all_ci[p, 'lower']), round(all_ci[p, 'upper']))
      )
  } else if (p == 'nu') {
    gg_tmp <- gg_tmp + 
      scale_x_continuous(breaks = c(0.9, 0.95, 1.0))
  }

  if (i %in% seq(1, 16, 4)) {
    h1_plots[[i]] <- gg_tmp
  } else {
    h1_plots[[i]] <- gg_tmp + theme(
      axis.text.y = element_blank(), axis.ticks.y = element_blank(),
      axis.title.y = element_blank()
      )
  }
}

cowplot::plot_grid(
  plotlist = h1_plots,
  align = 'h', ncol = 4,
  rel_widths = c(1.25, 1, 1, 1)
)
@
\caption{\label{m1Profs}MCAP confidence intervals for Model 1 parameters.}
\end{figure}

<<Load Model 2 Confidence Intervals, include=FALSE, message=FALSE, echo=FALSE>>=
h2_profiles <- readRDS("../model2/h2_profiles.rds") %>% as.data.frame()

prof_params <- c(
  "Mu",
  "BetaW",
  "v",
  "phase",
  "Beta",
  "sigma"
)

h2_plots <- list()
for (i in 1:length(prof_params)) {
  p = prof_params[i]
  if (p == 'Mu') {
    my_lab <- TeX('$\\mu_W')
  } else if (p == 'Beta') {
    my_lab <- TeX('$\\beta$')
  } else if (p == 'BetaW') {
    my_lab <- TeX('$\\beta_W$')
  } else if (p == 'v') {
    my_lab <- TeX("$\\psi")
  } else if (p == 'sigma') {
    my_lab <- TeX('$\\mu_{\\text{RS}}$')
  } else if (p == 'phase') {
    my_lab <- TeX('$\\phi')
  }

  if (p == 'Beta') {
    tmp_df_beta <- h2_profiles %>% 
      filter(prof_var == 'Beta') %>% 
      # filter(logLik >= -21965) %>%
      arrange(p)

    vals_beta <- tmp_df_beta[, 'Beta']^(1/5)

    profile_like_beta <- splinefun(x = vals_beta, y = tmp_df_beta[, 'logLik'])
    # max_profile <- optimize(profile_like, interval = range(vals), maximum = TRUE)
    max_profile_beta <- max(h2_profiles$logLik)
    threshold_beta <- max_profile_beta - 1.92
    plot_threshold_beta <- max_profile_beta - 6
    profile_roots_beta <- function(x) {profile_like_beta(x) - threshold_beta}
    plot_limits_beta <- function(x) {profile_like_beta(x) - plot_threshold_beta}

    upper_bound_beta <- uniroot(
      profile_roots_beta, 
      interval = c(vals_beta[which.max(tmp_df_beta$logLik)], max(vals_beta))
    )$root

    max_x_beta <- uniroot(
      plot_limits_beta, 
      interval = c(0, 7e-06^(1/5))
    )$root

    gg_tmp_beta <- ggplot() + 
      geom_point(
        data = tmp_df_beta %>% filter(prof_var == p) %>% filter(Beta <= max_x_beta^5),
        aes(x = Beta, y = logLik)
      ) + 
      geom_function(fun = function(x) profile_like_beta(x^(1/5))) + 
      geom_vline(xintercept = upper_bound_beta^5, linetype = 'dashed') + 
      geom_vline(xintercept = 0, col = 'blue') + 
      labs(y = 'Log-Likelihood', title = my_lab) + 
      theme(
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 10)
      ) +
      theme_bw() +
      scale_y_continuous(limits = c(max_profile_beta-6, max_profile_beta)) +
      scale_x_continuous(limits = c(0, max_x_beta^5)) +
      theme(
        axis.title.x = element_blank(), 
        plot.title = element_text(
          hjust = 0.5, size = 8, margin = margin(t = 0, unit = "pt")
        ),
        axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 7),
        axis.text.x = element_text(size = 6.5),
        plot.margin = margin(5.5, 3, 0, -1, unit = 'pt')
      )

    if (i %in% c(1, 4)) {
      h2_plots[['beta']] <- gg_tmp_beta
    } else {
      h2_plots[['beta']] <- gg_tmp_beta + theme(
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
      )
    }

  } else if (p == 'sigma') {
    tmp_df_sigma <- h2_profiles %>% 
      filter(prof_var == 'sigma') %>% 
      # filter(logLik >= -21965) %>%
      arrange(p)

    vals_sigma <- tmp_df_sigma[, 'sigma']^(1/5)

    profile_like_sigma <- splinefun(x = vals_sigma, y = tmp_df_sigma[, 'logLik'])
    # max_profile <- optimize(profile_like, interval = range(vals), maximum = TRUE)
    max_profile_sigma <- max(h2_profiles$logLik)
    threshold_sigma <- max_profile_sigma - 1.92
    plot_threshold_sigma <- max_profile_sigma - 6
    profile_roots_sigma <- function(x) {profile_like_sigma(x) - threshold_sigma}
    plot_limits_sigma <- function(x) {profile_like_sigma(x) - plot_threshold_sigma}

    upper_bound_sigma <- uniroot(
      profile_roots_sigma, 
      interval = c(vals_sigma[which.max(tmp_df_sigma$logLik)], max(vals_sigma))
    )$root

    max_x_sigma <- uniroot(
      plot_limits_sigma, 
      interval = c(0, 0.0022^(1/5))
    )$root

    gg_tmp_sigma <- ggplot() + 
      geom_point(
        data = tmp_df_sigma %>% filter(prof_var == 'sigma') %>% filter(sigma <= max_x_sigma^5),
        aes(x = sigma, y = logLik)
      ) + 
      geom_function(fun = function(x) profile_like_sigma(x^(1/5))) + 
      geom_vline(xintercept = upper_bound_sigma^5, linetype = 'dashed') + 
      geom_vline(xintercept = 0, col = 'blue') + 
      labs(y = 'Log-Likelihood', title = my_lab) + 
      theme(
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 10)
      ) +
      theme_bw() +
      scale_y_continuous(limits = c(max_profile_sigma-6, max_profile_sigma)) +
      scale_x_continuous(limits = c(0, max_x_sigma^5)) +
      theme(
        axis.title.x = element_blank(), 
        plot.title = element_text(
          hjust = 0.5, size = 8, margin = margin(t = 0, unit = "pt")
        ),
        axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 7),
        axis.text.x = element_text(size = 6.5),
        plot.margin = margin(5.5, 3, 0, -1, unit = 'pt')
      )

    if (i %in% c(1, 4)) {
      h2_plots[['sigma']] <- gg_tmp_sigma
    } else {
      h2_plots[['sigma']] <- gg_tmp_sigma + theme(
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
      )
    }

  } else {
    tmp_df <- h2_profiles %>% 
      filter(prof_var == p) %>% 
      # filter(logLik >= -21957.26 - 4) %>%
      arrange(p)

    vals <- tmp_df[, p]

    profile_like <- splinefun(x = tmp_df[, p], y = tmp_df[, 'logLik'])
    max_profile <- optimize(profile_like, interval = range(vals), maximum = TRUE)
    threshold <- max_profile$objective - 1.92
    plot_threshold <- max_profile$objective - 6
    profile_roots <- function(x) {profile_like(x) - threshold}
    plot_limits <- function(x) {profile_like(x) - plot_threshold}

    lower_bound <- uniroot(
      profile_roots, 
      interval = c(min(vals), max_profile$maximum)
    )$root

    upper_bound <- uniroot(
      profile_roots, 
      interval = c(vals[which.max(tmp_df$logLik)], max(vals))
    )$root

    add_piece <- c(
      'Mu' = 3000, 'BetaW' = 0.3, 'phase' = 0.2, v = 0.5 
    )

    max_x <- uniroot(
      plot_limits, 
      interval = c(max_profile$maximum, max(vals) + add_piece[p])
    )$root

    min_x <- uniroot(
      plot_limits, 
      interval = c(min(vals) - add_piece[p], max_profile$maximum)
    )$root

    gg_tmp <- ggplot() + 
      geom_point(
        data = h2_profiles %>% 
          filter(prof_var == p) %>% 
          filter(.data[[p]] <= max_x & .data[[p]] >= min_x) ,
        aes(x = .data[[p]], y = logLik)
      ) + 
      geom_function(fun = profile_like) + 
      geom_vline(xintercept = lower_bound, linetype = 'dashed') + 
      geom_vline(xintercept = upper_bound, linetype = 'dashed') + 
      geom_vline(xintercept = max_profile$maximum, col = 'blue') + 
      labs(y = 'Log-Likelihood', title = my_lab) + 
      theme(
        axis.text = element_text(size = 8),
        axis.title = element_text(size = 10)
      ) +
      theme_bw() +
      # scale_y_continuous(limits = c(-21963, -21957)) + 
      scale_y_continuous(limits = c(max(h2_profiles$logLik)-6, max(h2_profiles$logLik))) +
      scale_x_continuous(limits = c(min_x, max_x)) + 
      theme(
        axis.title.x = element_blank(), 
        plot.title = element_text(
          hjust = 0.5, size = 8, margin = margin(t = 0, unit = "pt")
        ),
        axis.title.y = element_text(size = 8),
        axis.text.y = element_text(size = 7),
        axis.text.x = element_text(size = 6.5),
        plot.margin = margin(5.5, 3, 0, -1, unit = 'pt')
      )

    if (i %in% c(1, 4)) {
      h2_plots[[p]] <- gg_tmp
    } else {
      h2_plots[[p]] <- gg_tmp + theme(
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
      )
    }
  }
}
@

\begin{figure}[ht]
<<m2Profs, fig.height=3, fig.width=5, echo=FALSE, warning=FALSE, message=FALSE>>=
cowplot::plot_grid(
  plotlist = h2_plots, ncol = 3,
  align = 'h',
  rel_widths = c(1.15, 1, 1)
)
@
\caption{\label{m2Profs}MCAP confidence intervals for Model 2 parameters.}
\end{figure}

<<Load Model 3 Profiles, include=FALSE, message=FALSE, echo=FALSE>>=
H3_profiles1 <- readRDS('../model3/run_level_3/h3_profiles.rds') %>% 
  filter(prof_var == 'std_W')
H3_profiles2 <- readRDS("../model3/run_level_3/h3_profiles2.rds")
H3_profiles3 <- readRDS("../model3/run_level_3/h3_profiles3.rds")

H3_MLE_evals <- readRDS('../model3/run_level_3/h3_MLE_eval.rds')

H3_profiles <- bind_rows(H3_profiles1, H3_profiles2, H3_profiles3)

prof_params <- unique(H3_profiles$prof_var)

H3_profiles_long <- H3_profiles %>%
  select(prof_var, logLik, logLik_se, all_of(paste0(prof_params, '1'))) %>%
  rename_with(~ sub("1$", "", .x)) %>%
  select(-epsilon, -XthetaA) %>% 
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(variable == prof_var) %>%
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 1)

H3_remaining_long <- H3_profiles %>% 
  select(prof_var, logLik, logLik_se, epsilon1, XthetaA1) %>% 
  rename(epsilon = epsilon1, XthetaA = XthetaA1) %>% 
  mutate(XthetaA = sqrt(XthetaA)) %>%
  mutate(epsilon = epsilon^2) %>%
  pivot_longer(
    cols = -c(prof_var, logLik, logLik_se),
    names_to = 'variable',
    values_to = 'value'
  ) %>%
  filter(variable == prof_var) %>%
  group_by(prof_var, value) %>%
  slice_max(order_by = logLik, n = 2)

H3_profiles_long <- bind_rows(
  H3_profiles_long, H3_remaining_long
)
@

<<Create Confidence intervals>>=
all_mcaps <- list()
all_ci <- list()
for (p in prof_params) {

  mcap_tmp <- mcap(
    logLik = H3_profiles_long %>% filter(prof_var == p) %>% pull(logLik),
    parameter = H3_profiles_long %>% filter(prof_var == p) %>% pull(value),
    span = case_when(
      p == 'r' ~ 1,
      p == 'XthetaA' ~ 0.7,
      TRUE ~ 0.75
    )
  )

  all_mcaps[[p]] <- mcap_tmp
  tmp_ci <- mcap_tmp$ci
  names(tmp_ci) <- c("lower", "upper")

  all_ci[[p]] <- c(tmp_ci, 'mle' = mcap_tmp$mle)
}

all_mcaps <- purrr::map_df(
  prof_params,
  function(x) all_mcaps[[x]]$fit %>% mutate(prof_var = x)
)

all_ci <- all_ci |> as.data.frame() |> t() |> as.data.frame()
all_ci$prof_var <- rownames(all_ci)

# Transform epsilong (rho) to natural scale
all_ci <- all_ci %>% 
  mutate(
    lower = case_when(
      prof_var == 'epsilon' ~ sqrt(lower), 
      # prof_var == 'XthetaA' ~ lower^2,
      TRUE ~ lower
    ),
    upper = case_when(
      prof_var == 'epsilon' ~ sqrt(upper), 
      # prof_var == 'XthetaA' ~ upper^2,
      TRUE ~ upper
    ),
    mle = case_when(
      prof_var == 'epsilon' ~ sqrt(mle), 
      # prof_var == 'XthetaA' ~ mle^2,
      TRUE ~ mle
    )
  )

all_mcaps <- all_mcaps %>% 
  mutate(
    parameter = case_when(
      prof_var == 'epsilon' ~ sqrt(parameter), 
      # prof_var == 'XthetaA' ~ parameter^2,
      TRUE ~ parameter
      )
  )

H3_profiles_long <- H3_profiles_long %>% 
  mutate(
    value = case_when(
      prof_var == 'epsilon'~ sqrt(value), 
      # prof_var == 'XthetaA' ~ value^2,
      TRUE ~ value
    )
  )

# This chunk of code returns the MLE for model 3
H3_MLE_evals %>%
  slice_max(order_by = logLik, n = 1) %>%
  select(logLik, all_of(paste0(prof_params, '1'))) %>%
  rename_with(~ sub("1$", "", .x)) %>%
  mutate(XthetaA = sqrt(XthetaA)) %>%
  pivot_longer(
    cols = -c(logLik),
    names_to = 'prof_var',
    values_to = 'value'
  ) -> mles
@

\begin{figure}[ht]
<<m3Profs, fig.height=4.6, fig.width=5.8, echo=FALSE, warning=FALSE, message=FALSE>>=
h3_plots <- list()
for (i in 1:length(prof_params)) {
  p = prof_params[i]
  if (p == 'mu_B') {
    my_lab <- TeX('$\\zeta')
  } else if (p == 'XthetaA') {
    my_lab <- TeX('$\\sqrt{\\epsilon_W}$')
  } else if (p == 'thetaI') {
    my_lab <- TeX('$\\mu_W$')
  } else if (p == 'lambdaR') {
    my_lab <- TeX('$a$')
  } else if (p == 'r') {
    my_lab <- TeX('$r$')
  } else if (p == 'std_W') {
    my_lab <- TeX('$\\sigma$')
  } else if (p == 'epsilon') {
    my_lab <- TeX('$\\rho$')
  } else if (p == 'k') {
    my_lab <- TeX('$\\psi$')
  }

  gg_tmp <- ggplot() +
    geom_point(
      data = H3_profiles_long %>% filter(prof_var == p), 
      aes(x = value, y = logLik)
    ) +
    geom_line(
      data = all_mcaps %>% filter(prof_var == p), 
      aes(x = parameter, y = smoothed), 
      col = 'blue'
    ) +
    # geom_line(data = all_mcaps, aes(x = parameter, y = quadratic), col = 'red') +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = lower), linetype = 'dashed'
    ) +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = upper), 
      linetype = 'dashed'
    ) +
    geom_vline(
      data = all_ci %>% filter(prof_var == p), 
      aes(xintercept = mle), col = 'blue'
    ) +
    # This commented out code below will plot the MLE along with the confidence 
    # intervals for the given parameter. 
    # geom_point(
    #   data = mles %>% filter(prof_var == p),
    #   aes(x = value, y = logLik), col = 'red'
    # ) +
    labs(y = 'Log-Likelihood', title = my_lab) +
    theme(
      axis.text = element_text(size = 8),
      axis.title = element_text(size = 10)
    ) +
    theme_bw() +
    scale_y_continuous(limits = c(-17450, max(H3_profiles$logLik))) +
    # scale_y_continuous(limits = c(-17450, max(H3_MLE_evals$logLik))) +
    theme(
      axis.title.x = element_blank(), 
      plot.title = element_text(
        hjust = 0.5, size = 8, margin = margin(t = 0, unit = "pt")
        ),
      axis.title.y = element_text(size = 8),
      axis.text.y = element_text(size = 7),
      axis.text.x = element_text(size = 6.5),
      plot.margin = margin(5.5, 3, 0, -1, unit = 'pt')
    ) + 
    theme(legend.position = 'none')

  if (i %in% seq(1, 9, 3)) {
    h3_plots[[i]] <- gg_tmp
  } else {
    h3_plots[[i]] <- gg_tmp + theme(
      axis.text.y = element_blank(), axis.ticks.y = element_blank(),
      axis.title.y = element_blank()
      )
  }
}

cowplot::plot_grid(
  plotlist = h3_plots,
  align = 'h', ncol = 3,
  rel_widths = c(1.25, 1, 1, 1)
)
@
\caption{\label{m3Profs}MCAP confidence intervals for Model 1 parameters.}
\end{figure}

%%%%%%%%%% END

\bibliography{../bib-haiti}

\end{document}